%%!TEX root = ./UserManual.tex
\chapter{Code}
\label{chap:code}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% GenGeoPop
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{GenGeoPop}
\label{section:gengeopop}

\subsection{Background}
\label{subsection:background}

To explain the algorithms used for generating the geography of the countries and their respective population, we have to introduce some background concepts:

\begin{description}
    \item[ContactPool]:
        A pool of persons that potentially come in contact with each other.
        If they actually do and if a transmission of an infection occurs depends on the algorithms in the simulator.
        We distinguish different types of ContactPools associated with work, school, community and households.
        Note that there is a difference between the size of a school (for example 500 students) and the size of the ContactPools inside the school (20 students).
    \item[ContactCenter]:
        A group of one or more ContactPools of the same type.
        This allows for a single College to contain multiple classes for example.
        It is also used to contain a single ContactPool, for example in the case of a Household.
    \item[K-12 student]: Part of the population from 3 until 18 years of age that are obligated (at least in Belgium) to attend a school. Students that skip or repeat years are not accounted for.
    \item[College student]:
        Persons older than 18 and younger than 26 years of age that attend a higher education. For simplicity we will group all forms of higher eduction into the same type of ContactCenter, a College. A fraction of college students will attend a college ``close to home'' and the others will attend a college ``far from home''.
        Most educations don't last for 8 years, but using this number we compensate for changes in the field of study, doctoral studies, advanced masters and repeating a failed year of study.
    \item[Employable]:
        We consider people of ages 18 to 65 as employable. A fraction of people between 18 and 26 will attend a college, and the complementary fraction will be employable.
    \item[Active population]:
        The fraction of the employable part of the population that is actually working. A fraction of these workers will work ``close to home'' and the complementary fraction will commute to a workplace further away.
    \item[Household profile]:
        The composition of households in terms of the number of members and their age is an important factor in the simulation. In this case the profile is not defined by the age of its members or fractions, but through a set of reference households. This set contains a sample of households which are representative of the whole population in their composition.
    \item[GeoGrid locations]:
        Our data only allows for a very limited geographical resolution. We have the longitude and latitude of cities and municipalities (a distinction we will not make), which we shall use to create GeoGrid locations for the domicile of the households.
        All households in the same location are mapped to the coordinates of the location's center. These locations with corresponding coordinates will form a grid that will contain the complete simulation area.
\end{description}

\subsection{Generating the geography}
\label{subsection:gengeo}
We start by generating the geographical component, we call this a GeoGrid.
This contains locations with an id, name, province, coordinates and nominative population figures.
These locations contain multiple ContactCenters, like Schools and Households, which in turn contain ContactPools.
This structure is internally generated by several ``Generators" and will afterwards be used by ``Populators" to fill the ContactPools.
We construct the following types of ContactCenter:

\begin{description}
    \item[K-12 Schools]:
        These schools contain on average 500 students, with an average of 20 students per class, which corresponds to 25 ContactPools.
        The total amount of schools in the region is determined by the size of the population and the fraction of those people who attend a K-12 school.
        These generated schools are then assigned to a location by a draw from a discrete distribution based on the relative population size in each location.
    \item[Colleges]:
        These contain an average of 3000 students and 150 students per ContactPool.
        They are exclusively assigned to the 10 biggest locations in terms of population size.
        Within those 10 locations we again use a discrete distribution based on the relative population of the city compared to the total population in these top 10 cities.
    \item[Workplaces]:
        The assignment of workplaces is analogous to that of K-12 schools, but now we account for commuting information to determine the actual number of workers in a location.
        We find this amount by the working people who live there plus the people that commute to this location minus the ones that live there but commute away.
        A workplace contains on average 20 people.
    \item[Communities]:
        We create both primary and secondary communities, each containing 2000 persons from all ages and occupations.
        The assignement is analogous to that of K-12 schools.
    \item[Households]:
        The number of households is determined by the average size of a household in the reference profile and the total population size.
        The assignement is again analogous to that of K-12 schools.
\end{description}

\subsection{Generating the population}
\label{subsection:genpop}
TODO

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Travel
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Travel}
\label{section:travel}
TODO
